Ionospheric Disturbances in Mexican Territory Produced by Objects Entering the Athmosphere from Space \tnoteref{tnote1}

### 2021-06-19 paper/manuscript.tex
Improve vTeC code, insert Kp index graphs and work with paper
- Update Journal
- Insert Kp index graphs
- Normalize colormap in vTEC maps
- Few advances in paper
- Working in make vTEC maps for all events

%% main text
\section{Introduction}
\label{sec1}
Please use \verb+elsarticle.cls+ for typesetting your paper. Additionally,
make sure not to remove the package \verb+jasr.sty+ already included in the
preamble:
%\begin{verbatim} 
%  \usepackage{jasr}
%\end{verbatim}

Make sure to have the file \verb+model5-names.bst+ to produce the references in
the correct format.

Any instructions relevant to the \verb+elsarticle.cls+ are
applicable here as well. See the online instruction available at:
\makeatletter
\if@twocolumn
\begin{verbatim}
 https://support.stmdocs.in/wiki/
 index.php?title=Elsarticle.cls
\end{verbatim}
\else
\begin{verbatim}
 https://support.stmdocs.in/wiki/index.php?title=Elsarticle.cls
 \end{verbatim}
\fi
\makeatother

Following commands are defined for this journal which are not in
\verb+elsarticle.cls+. 
\begin{verbatim}
  \received{}
  \finalform{}
  \accepted{}
  \availableonline{}
  \communicated{}
\end{verbatim}

\subsection{Subsection}
This is only a \LaTeX\ template if you need one. 
See the detailed guidelines for manuscript preparation 
and submission at:
\begin{verbatim}
https://ees.elsevier.com/asr
\end{verbatim}

@@

We selected a sample of meteors which were observed in mexican territory from the Geostationary Lightning Mapper \citep{GOODMAN:2013}. The sample was selected in such way we chose the most probable obects to be detected by GPS sources in mexican territory:

@@

 For the selected sample, shown in , we obtained RINEX data from the TLALOCNet and UNAVCO network databases \citep{Cabral-Cano:2018} to study potential alterations in the ionospehre due to the presence of the passing meteor in the day the meteor was reported. The list of the sample meteors is shown in table \ref{tab:table-meteors}.

\begin{table*}
\centering
\caption{List of meteors passing through Mexico.}
\label{tab:table-meteors}
\begin{tabular}{|r|r|r|r|r|r|}
\hline
ID & Fecha & Start Time (UT)  & Duration (seconds) & Latitud & Longitud \\
\hline
01 & 23-05-2019 & 16:36:18 & $0.197\pm 0.0000$ & $24.30 \pm 0.000$ & $-101.60 \pm 0.849$\\
02 & 18-07-2019 & 14:30:30 & $0.058\pm 0.0000$ & $27.20 \pm 0.000$ & $-103.15 \pm 0.778$\\
03 & 10-08-2019 & 11:18:48 & $0.199\pm 0.0757$ & $21.50 \pm 0.000$ & $-102.50  \pm 0.849$\\
04 & 03-10-2019 & 07:55:33 & $0.106\pm 0.0297$ & $25.65 \pm 0.071$ & $-96.25 \pm   0.778$\\
05 & 09-10-2019 & 06:08:11 & $0.103\pm 0.0078$ & $23.60 \pm 0.000$ & $-111.95 \pm  0.212$\\
06 & 12-10-2019 & 11:12:27 & $0.006\pm 0.0000$ & $28.8 \pm 0.000$ & $-111.3 \pm 0.000$\\
07 & 16-11-2019 & 09:36:04 & $0.396\pm 0.0134$ & $20.30 \pm 0.000$ & $-100.55 \pm  0.919$\\
08 & 17-11-2019 & 15:36:01 & $0.116\pm 0.0035$ & $31.70 \pm 0.000$ & $-117.70 \pm  1.131$\\
09 & 19-11-2019 & 07:57:40 & $0.097\pm 0.1138$ & $20.00 \pm 0.000$ & $-88.40 \pm  1.131$\\
10 & 26-11-2019 & 13:23:20 & $0.078\pm 0.0290$ & $23.90 \pm 0.000$ & $-108.70 \pm  0.849$\\
11 & 04-12-2019 & 09:42:54 & $0.173\pm 0.0028$ & $31.50 \pm 0.000$ & $-113.65 \pm  0.919$\\
12 & 15-12-2019 & 14:50:49 & $0.127\pm 0.0134$ & $27.70 \pm 0.000$ & $-114.10 \pm  0.849$\\
13 & 29-12-2019 & 16:16:35 & $0.062\pm 0.0134$ & $29.60 \pm 0.000$ & $-116.35 \pm  0.919$\\
14 & 03-01-2020 & 14:10:17 & $0.113\pm 0.0085$ & $30.20 \pm 0.000$ & $-117.65 \pm  0.919$\\
15 & 06-01-2020 & 16:39:27 & $0.118\pm 0.0042$ & $31.40 \pm 0.000$ & $-108.20 \pm  0.990$\\
16 & 15-01-2020 & 15:00:33 & $0.213\pm 0.1351$ & $19.45 \pm 0.071$ & $-95.55 \pm   0.919$\\
17 & 12-02-2020 & 09:25:40 & $0.210\pm 0.0226$ & $18.90 \pm 0.000$ & $-93.50 \pm   0.849$\\
18 & 03-03-2020 & 12:33:27 & $0.062\pm 0.0007$ & $18.25 \pm 0.071$ & $-106.35 \pm 0.636$\\
19 & 31-03-2020 & 19:31:52 & $0.105\pm 0.0573$ & $28.45 \pm 0.071$ & $-112.05 \pm  0.636$\\
20 & 08-04-2020 & 16:25:28 & $0.120\pm 0.0926$ & $26.10 \pm 0.000$ & $-93.90 \pm   0.849$\\
21 & 18-04-2020 & 17:43:25 & $0.139\pm 0.0106$ & $29.00 \pm 0.000$ & $-106.55 \pm  0.919$\\
22 & 20-04-2020 & 16:05:22 & $0.318\pm 0.1655$ & $28.15 \pm 0.071$ & $-97.85 \pm   1.061$\\
23 & 25-04-2020 & 11:03:09 & $0.323\pm 0.0813$ & $32.15 \pm 0.071$ & $-111.60 \pm  1.131$\\
24 & 28-04-2020 & 19:31:52 & $0.105\pm 0.0573$ & $28.45 \pm 0.071$ & $-112.05 \pm  0.636$\\
25 & 08-05-2020 & 10:06:16 & $0.490\pm 0.0750$ & $21.60 \pm 0.000$ & $-92.40 \pm   0.849$\\
26 & 15-07-2020 & 19:58:28 & $0.693\pm 0.0495$ & $24.00 \pm 0.000$ & $-108.35 \pm  0.495$\\
27 & 07-08-2020 & 13:29:57 & $0.163\pm 0.0057$ & $28.80 \pm 0.000$ & $-106.05 \pm  0.919$\\
28 & 13-09-2020 & 16:41:59 & $0.184\pm 0.0078$ & $28.45 \pm 0.071$ & $-113.75 \pm  0.919$\\
29 & 30-09-2020 & 12:28:11 & $0.100\pm 0.0078$ & $24.90 \pm 0.000$ & $-110.90 \pm  0.849$\\
30 & 16-11-2020 & 12:28:11 & $0.100\pm 0.0078$ & $24.90 \pm 0.000$ & $-110.90 \pm  0.849$\\
31 & 17-11-2020 & 12:53:41 & $0.404\pm 0.0262$ & $23.00 \pm 0.000$ & $-102.45 \pm  0.919$\\
32 & 19-12-2020 & 10:18:14 & $0.407\pm 0.0110$ & $21.95 \pm 0.071$ & $-101.60 \pm  0.990$\\
33 & 23-12-2020 & 09:43:01 & $0.148\pm 0.0014$ & $25.75 \pm 0.071$ & $-111.25 \pm  0.778$\\
34 & 29-12-2020 & 15:20:54 & $0.118\pm 0.0014$ & $16.80 \pm 0.000$ & $-102.20 \pm  0.707$\\
35 & 31-03-2021 & 09:01:17 & $0.753\pm 0.3083$ & $20.15 \pm 0.071$ & $-92.95 \pm  0.212$\\\hline
\end{tabular}
\end{table*}

%\begin{table*}
%\centering
%\caption{List of meteors passing through Mexico.}
%\label{tab:table-meteors}
%\begin{tabular}{|l|l|l|l|l|l|}
%\hline
%\multirow{2}{2.5cm}{Date (yyyy-mm-dd)} & \multirow{2}{*}{Time (UT)} & \multirow{2}{*}{Latitude} & %\multirow{2}{*}{Longitude} & \multirow{2}{*}{$\Delta t$ (seconds)}   & \multirow{2}{*}{Location}\\
% & & & &  & \\
%\hline
%2015-07-19 & 07:06:26 & 20.6 & -87.6 & 7.5 & Quintana Roo, Mx. \\
%\hline
%2017-07-31 & 22:01:34 & 24.8 & -118.6 & 1.1 & Pacific Ocean, Near Baja California, Mx. \\
%\hline
%2017-10-01 & 12:43:53 & 36.2 & -111.4 & 0.1 & Arizona, USA \\
%\hline
%2017-11-15 & 03:26:46 & 34.2 & -111.4 & 0.08 & Arizona, USA \\
%\hline
%2019-02-01 & 18:17:10 & 22.5 & -83.8 & 2.3 & La Habana, Cuba \\
%\hline
%2020-02-19 & 2:10 & 19.68 & -101.2 & 3.5 & Michoac\'an, Mx. \\
%\hline
%2020-04-28 & 05:43:17 & 19.9 & -110.2 & 0.8 &  Pacific Ocean, near Colima, Mx. \\
%\hline
%2020-10-17 & 03:14:23 & 26.1 & -100.1 & 0.04 & Nuevo Le\'on, Mx. \\
%\hline 
%2020-11-20 & 2:57:10 & 26.5 & -98.7 & 0.01 & Tamaulipas, Mx.\\
%\hline
%\end{tabular}
%\end{table*}

@@

\section{Ionospheric background}

\section{How to include a figure?}
Change the file name of the figure below with your own 
figure. And remember to change the figure caption too!
\begin{verbatim}
\begin{figure}
  \centering
  \includegraphics[scale=0.5]{fig01}
  \caption{Write the figure caption here.}
  \label{fig:pendulum}
\end{figure}
\end{verbatim}

@@

\subsection{And a table?}
Just replace the text/values in the template table below 
with your own. You can change the number of 
lines/rows as necessary.

@@

\subsection{Equations}
Conventionally, in mathematical equations, variables and
anything that represents a value appear in italics. 
All equations should be numbered for easy referencing. 
The number should appear at the right margin.
\begin{eqnarray}
S'_{\mathrm{pg}} = \frac{S_{\mathrm{pg}} - \mathrm{min}(S_{\mathrm{pG}})}
  {\mathrm{max}(S_{\mathrm{pG}} - \mathrm{min}(S_{\mathrm{pG}}))}
\end{eqnarray}
In mathematical expressions 
in running text "/" should be used
for division (not a horizontal line).

\section{Citations}
%Citations in the text can be made using\\[6pt]
%\verb+\citet{NewmanGirvan2004}+\\[6pt]
%for citation in running text like in 
%\citet{NewmanGirvan2004} or using\\[6pt]
%\verb+\citep{Vehlowetal2013,NewmanGirvan2004}+\\[6pt]
%for citation within parentheses like in 
%\citep{Vehlowetal2013,NewmanGirvan2004}.

@@

\section{Reference style}
You can include the references in the main text file in \LaTeX
format. Alternately, you can include a separate bibliography
file (refs.bib in this example) and run the following set of 
commands:
\begin{verbatim}
==================

pdflatex myfile.tex

bibtex myfile (no extension in this line!)

pdflatex myfile.tex

pdflatex myfile.tex

==================
\end{verbatim}

\section{A sample entry in the bibliography file}
%{\fontsize{7.5pt}{9.6pt}\selectfont
\begin{verbatim}
==================

@ARTICLE{NewmanGirvan2004,
  author  = {Newman, M. E. J. and Girvan, M.},
  title   = {Finding and evaluating community 
               structure in networks},
  journal = {Phys. Rev. E.},
  volume  = {69},
  number  = {21},
  year    = {2004},
  pages   = {026113}
}

==================
\end{verbatim}
%}

>>>>

%% main text
\section{Introduction}
\label{sec1}
%Please use \verb+elsarticle.cls+ for typesetting your paper. Additionally,
%make sure not to remove the package \verb+jasr.sty+ already included in the
%preamble:
%\begin{verbatim} 
%  \usepackage{jasr}
%\end{verbatim}

%Make sure to have the file \verb+model5-names.bst+ to produce the references in
%the correct format.

%Any instructions relevant to the \verb+elsarticle.cls+ are
%applicable here as well. See the online instruction available at:
%\makeatletter
%\if@twocolumn
%\begin{verbatim}
% https://support.stmdocs.in/wiki/
% index.php?title=Elsarticle.cls
%\end{verbatim}
%\else
%\begin{verbatim}
% https://support.stmdocs.in/wiki/index.php?title=Elsarticle.cls
% \end{verbatim}
%\fi
%\makeatother

%Following commands are defined for this journal which are not in
%\verb+elsarticle.cls+. 
%\begin{verbatim}
%  \received{}
%  \finalform{}
%  \accepted{}
%  \availableonline{}
%  \communicated{}
%\end{verbatim}

%\subsection{Subsection}
%This is only a \LaTeX\ template if you need one. 
%See the detailed guidelines for manuscript preparation 
%and submission at:
%\begin{verbatim}
%https://ees.elsevier.com/asr
%\end{verbatim}

@@

We selected a sample of meteors which were observed in mexican territory from the Geostationary Lightning Mapper \citep{GOODMAN:2013}. We used the interactive database available at \url{https://neo-bolide.ndc.nasa.gov/#/} to get the events positions presented in this section. The sample was selected in such way we chose the most probable objects to be detected by GPS sources in mexican territory:

@@

 For the selected sample, we obtained RINEX data from the TlalocNet \citep{Cabral-Cano:2018} and UNAVCO network databases to study potential alterations in the ionospehre due to the presence of the passing meteor at the day the meteor was reported. The list of the sample meteors is shown in table \ref{tab:table-meteors}. The events are in chronological order. The reported duration, latitude and longitude correspond to the mean between measurements from satellites GOES-16 and GOES 17; in the same way, the uncertainties correspond to the standard deviation. Also their respective positions are available in figure \ref{fig:meteors-map}, where each label correspond to the ID (first column) of table \ref{tab:table-meteors}.

\begin{table*}
  \centering
  \caption{List of meteors passing through Mexico. The events are listed in chronological order. The listed duration, latitude and longitude correspond to the mean of the measurements of both GOES satellites. The uncertainties correspond to the respecting mean deviation.}
\label{tab:table-meteors}
\begin{tabular}{|r|r|r|r|r|r|}
\hline
ID & Date of event & Start Time (UT)  & Duration (seconds) & Latitude (deg) & Longitude (deg)\\
\hline
01 & 2019-05-23 & 16:36:18 & $0.197\pm 0.0000$ & $24.30 \pm 0.000$ & $-101.60 \pm 0.849$\\
02 & 2019-07-18 & 14:30:30 & $0.058\pm 0.0000$ & $27.20 \pm 0.000$ & $-103.15 \pm 0.778$\\
03 & 2019-08-10 & 11:18:48 & $0.199\pm 0.0757$ & $21.50 \pm 0.000$ & $-102.50  \pm 0.849$\\
04 & 2019-10-03 & 07:55:33 & $0.106\pm 0.0297$ & $25.65 \pm 0.071$ & $-96.25 \pm   0.778$\\
05 & 2019-10-09 & 06:08:11 & $0.103\pm 0.0078$ & $23.60 \pm 0.000$ & $-111.95 \pm  0.212$\\
06 & 2019-10-12 & 11:12:27 & $0.006\pm 0.0000$ & $28.8 \pm 0.000$ & $-111.3 \pm 0.000$\\
07 & 2019-11-16 & 09:36:04 & $0.396\pm 0.0134$ & $20.30 \pm 0.000$ & $-100.55 \pm  0.919$\\
08 & 2019-11-17 & 15:36:01 & $0.116\pm 0.0035$ & $31.70 \pm 0.000$ & $-117.70 \pm  1.131$\\
09 & 2019-11-19 & 07:57:40 & $0.097\pm 0.1138$ & $20.00 \pm 0.000$ & $-88.40 \pm  1.131$\\
10 & 2019-11-26 & 13:23:20 & $0.078\pm 0.0290$ & $23.90 \pm 0.000$ & $-108.70 \pm  0.849$\\
11 & 2019-12-04 & 09:42:54 & $0.173\pm 0.0028$ & $31.50 \pm 0.000$ & $-113.65 \pm  0.919$\\
12 & 2019-12-15 & 14:50:49 & $0.127\pm 0.0134$ & $27.70 \pm 0.000$ & $-114.10 \pm  0.849$\\
13 & 2019-12-29 & 16:16:35 & $0.062\pm 0.0134$ & $29.60 \pm 0.000$ & $-116.35 \pm  0.919$\\
14 & 2020-01-03 & 14:10:17 & $0.113\pm 0.0085$ & $30.20 \pm 0.000$ & $-117.65 \pm  0.919$\\
15 & 2020-01-06 & 16:39:27 & $0.118\pm 0.0042$ & $31.40 \pm 0.000$ & $-108.20 \pm  0.990$\\
16 & 2020-01-15 & 15:00:33 & $0.213\pm 0.1351$ & $19.45 \pm 0.071$ & $-95.55 \pm   0.919$\\
17 & 2020-02-12 & 09:25:40 & $0.210\pm 0.0226$ & $18.90 \pm 0.000$ & $-93.50 \pm   0.849$\\
18 & 2020-03-03 & 12:33:27 & $0.062\pm 0.0007$ & $18.25 \pm 0.071$ & $-106.35 \pm 0.636$\\
19 & 2020-03-31 & 19:31:52 & $0.105\pm 0.0573$ & $28.45 \pm 0.071$ & $-112.05 \pm  0.636$\\
20 & 2020-04-08 & 16:25:28 & $0.120\pm 0.0926$ & $26.10 \pm 0.000$ & $-93.90 \pm   0.849$\\
21 & 2020-04-18 & 17:43:25 & $0.139\pm 0.0106$ & $29.00 \pm 0.000$ & $-106.55 \pm  0.919$\\
22 & 2020-04-20 & 16:05:22 & $0.318\pm 0.1655$ & $28.15 \pm 0.071$ & $-97.85 \pm   1.061$\\
23 & 2020-04-25 & 11:03:09 & $0.323\pm 0.0813$ & $32.15 \pm 0.071$ & $-111.60 \pm  1.131$\\
24 & 2020-04-28 & 19:31:52 & $0.105\pm 0.0573$ & $28.45 \pm 0.071$ & $-112.05 \pm  0.636$\\
25 & 2020-05-08 & 10:06:16 & $0.490\pm 0.0750$ & $21.60 \pm 0.000$ & $-92.40 \pm   0.849$\\
26 & 2020-07-15 & 19:58:28 & $0.693\pm 0.0495$ & $24.00 \pm 0.000$ & $-108.35 \pm  0.495$\\
27 & 2020-08-07 & 13:29:57 & $0.163\pm 0.0057$ & $28.80 \pm 0.000$ & $-106.05 \pm  0.919$\\
28 & 2020-09-13 & 16:41:59 & $0.184\pm 0.0078$ & $28.45 \pm 0.071$ & $-113.75 \pm  0.919$\\
29 & 2020-09-30 & 12:28:11 & $0.100\pm 0.0078$ & $24.90 \pm 0.000$ & $-110.90 \pm  0.849$\\
30 & 2020-11-16 & 12:28:11 & $0.100\pm 0.0078$ & $24.90 \pm 0.000$ & $-110.90 \pm  0.849$\\
31 & 2020-11-17 & 12:53:41 & $0.404\pm 0.0262$ & $23.00 \pm 0.000$ & $-102.45 \pm  0.919$\\
32 & 2020-12-19 & 10:18:14 & $0.407\pm 0.0110$ & $21.95 \pm 0.071$ & $-101.60 \pm  0.990$\\
33 & 2020-12-23 & 09:43:01 & $0.148\pm 0.0014$ & $25.75 \pm 0.071$ & $-111.25 \pm  0.778$\\
34 & 2020-12-29 & 15:20:54 & $0.118\pm 0.0014$ & $16.80 \pm 0.000$ & $-102.20 \pm  0.707$\\
35 & 2021-03-31 & 09:01:17 & $0.753\pm 0.3083$ & $20.15 \pm 0.071$ & $-92.95 \pm  0.212$\\\hline
\end{tabular}
\end{table*}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{../meteors_map}
  \caption{Positions of events from table \ref{tab:table-meteors}. The label of each point correspond to the ID (first column) of the referred table.}
  \label{fig:meteors-map}
\end{figure}

@@

\section{Ionospheric background and vTEC maps}

\section{Frecuency Analysis}

\section{Discussion}

%\section{How to include a figure?}
%Change the file name of the figure below with your own 
%figure. And remember to change the figure caption too!
%\begin{verbatim}
%\begin{figure}
%  \centering
%  \includegraphics[scale=0.5]{fig01}
%  \caption{Write the figure caption here.}
%  \label{fig:pendulum}
%\end{figure}
%\end{verbatim}

@@

%\subsection{And a table?}
%Just replace the text/values in the template table below 
%with your own. You can change the number of 
%lines/rows as necessary.

@@

%\subsection{Equations}
%Conventionally, in mathematical equations, variables and
%anything that represents a value appear in italics. 
%All equations should be numbered for easy referencing. 
%The number should appear at the right margin.
%\begin{eqnarray}
%S'_{\mathrm{pg}} = \frac{S_{\mathrm{pg}} - \mathrm{min}(S_{\mathrm{pG}})}
%  {\mathrm{max}(S_{\mathrm{pG}} - \mathrm{min}(S_{\mathrm{pG}}))}
%\end{eqnarray}
%In mathematical expressions 
%in running text "/" should be used
%for division (not a horizontal line).

%\section{Citations}
%Citations in the text can be made using\\[6pt]
%\verb+\citet{NewmanGirvan2004}+\\[6pt]
%for citation in running text like in 
%\citet{NewmanGirvan2004} or using\\[6pt]
%\verb+\citep{Vehlowetal2013,NewmanGirvan2004}+\\[6pt]
%for citation within parentheses like in 
%\citep{Vehlowetal2013,NewmanGirvan2004}.

@@

%\section{Reference style}
%You can include the references in the main text file in \LaTeX
%format. Alternately, you can include a separate bibliography
%file (refs.bib in this example) and run the following set of 
%commands:
%\begin{verbatim}
%==================

%pdflatex myfile.tex

%bibtex myfile (no extension in this line!)

%pdflatex myfile.tex

%pdflatex myfile.tex

%==================
%\end{verbatim}

%\section{A sample entry in the bibliography file}
%{\fontsize{7.5pt}{9.6pt}\selectfont
%\begin{verbatim}
%==================

%@ARTICLE{NewmanGirvan2004,
%  author  = {Newman, M. E. J. and Girvan, M.},
%  title   = {Finding and evaluating community 
%               structure in networks},
%  journal = {Phys. Rev. E.},
%  volume  = {69},
%  number  = {21},
%  year    = {2004},
%  pages   = {026113}
%}

%==================
%\end{verbatim}
%}
